Ages of Young Stars and Planets in the Kepler Field

### 2024-06-07 paper/ms.tex
update ms

Sub-gigayear planets can be particularly informative for studies of
planet evolution.  Our catalog has \nplyounggyro\ confirmed and
candidate planets with median ages below 1\,Gyr.  Requiring $t_{\rm
gyro}$$<$1\,Gyr at 2$\sigma$ yields \nplyounggyrotwosigma\ planets
orbiting \nplhostsyounggyrotwosigma\ stars.  The youth claim is most
secure for the subsets that are either in clusters, or that have
independent rotation and lithium-based ages.

@@

\subsubsection{Rotation vs.~Lithium Ages}

@@

\subsubsection{New Young Planets}

@@

Highlights from these systems include Kepler-1529, Kepler-1565,
Kepler-1312, and Kepler-1629.  Kepler-1529b is a
$\approx$100$\pm$50\,Myr mini-Neptune with well-measured rotation and
lithium ages.  Kepler-1565b is an analogous $\approx$170-230\,Myr
super-Earth.  Kepler-1312 is a $t_{\rm gyro}$=\kepthirteentwelve\,Myr
near-Solar analog with an Earth-sized planet on a one-day orbit, and a
mini-Neptune on a five-day orbit.  The Earth-sized planet,
Kepler-1312~c, along with the near-identical Kepler-1561~b ($t_{\rm
gyro}$=\kepfifteensixone\,Myr), rank among the youngest Earth-sized
planets currently known, comparable to planets such as HD~63433d
\citep[1.1\,$R_\oplus$, $414$$\pm$$23$\,Myr;][]{2024AJ....167...54C}
and TOI-1807~b \citep[1.3\,$R_\oplus$,
180$\pm$40\,Myr;][]{2021AJ....162...54H}.  Kepler-1629~b, with a size
just two-thirds that of Earth (0.67$\pm$0.05\,$R_\oplus$), is
marginally older, also based on rotation (\kepsixteentwonine\,Myr).
The lithium age from a reconnaissance TRES spectrum confirms the
point.

>>>>

Sub-gigayear planets can be particularly informative for studies of
planet evolution.  Our catalog has \nplyounggyro\ confirmed and
candidate planets with median ages below 1\,Gyr.  Requiring $t_{\rm
gyro}$$<$1\,Gyr at 2$\sigma$ yields \nplyounggyrotwosigma\ planets
orbiting \nplhostsyounggyrotwosigma\ stars.  The youth claim is most
secure for the systems that are either in clusters (Section~\ref{subsec:clusterplanets}), or that have
independent rotation and lithium-based ages (Section~\ref{subsec:rotnlithiumcomp}).
These samples provide broader context for the planets around field
stars that push the boundaries of the current young exoplanet census
(Section~\ref{subsec:notables}).

@@

\subsubsection{Rotation vs.~Lithium Ages}
\label{subsec:rotnlithiumcomp}

@@

\subsubsection{Notable New Young Planets}
\label{subsec:notables}

@@

Among these systems, four highlights are Kepler-1529, Kepler-1565,
Kepler-1312, and Kepler-1629.  Kepler-1529b is a
$\approx$100$\pm$50\,Myr mini-Neptune with well-measured rotation and
lithium ages.  Kepler-1565b is an analogous $\approx$170-230\,Myr
super-Earth.  Kepler-1312 is a $t_{\rm gyro}$=\kepthirteentwelve\,Myr
near-Solar analog with an Earth-sized planet on a one-day orbit, and a
mini-Neptune on a five-day orbit.  The Earth-sized planet,
Kepler-1312~c, along with the near-identical Kepler-1561~b ($t_{\rm
gyro}$=\kepfifteensixone\,Myr), rank among the youngest Earth-sized
planets currently known, comparable to planets such as HD~63433d
\citep[1.1\,$R_\oplus$, $414$$\pm$$23$\,Myr;][]{2024AJ....167...54C}
and TOI-1807~b \citep[1.3\,$R_\oplus$,
180$\pm$40\,Myr;][]{2021AJ....162...54H}.  Kepler-1629~b, with a size
just two-thirds that of Earth (0.67$\pm$0.05\,$R_\oplus$), is
marginally older, also based on rotation (\kepsixteentwonine\,Myr).
The lithium age from a reconnaissance TRES spectrum confirms the
point.
